Collaborative Visual Analysis on the Web with RCloud

### 2015-06-16 relatedwork.tex
fix typo in figure caption
yes

Broadly, the entire field of information visualization and visual analytics
revolves around improving how users solve problems involving data.
In this section, we focus specifically on
system work relating to problem-solving environments and multiple users.

\paragraph*{Social Data exploration and Analysis}
ManyEyes~\cite{Viegas:2007:MAS} was a landmark system designed
for the crowdsourced creation and publication of data
visualizations. Although ManyEyes only supported a limited number of
visual encodings, the system's success was both a precursor to more
sophisticated solutions and an early indication that the world-wide
web was a suitable platform for data visualization. One of the main
challenges we faced in designing RCloud was providing an experience
for \emph{consuming} visualizations as seamless as ManyEyes's, while
sacrificing as little as possible on the generality of the analyses
themselves.

@@

\paragraph*{Provenance and versioning} As
mentioned, one central issue in exploratory analysis is that
problems change quickly over time, often in the course of developing a
solution. As a result, systems should provide adequate support for
tracking \emph{changes} of the data analysis scripts. VisTrails was
one of the original systems for managing \emph{process provenance} ,
and demonstrated the value of capturing aspects of the processes that
surround data analysis experiments and tools, including detailed
history, collaboration, and deployment~\cite{Callahan:2006:VVM}.

@@

%% publishing web content such as knitR, Rpubs,
%% rCharts, and gg2v. These tools are not collaborative, but they are
%% aimed at better graphics and web applications.
%% . {\bf iPython
%%   notebooks (Jupyter)} are the closest work we are aware of.  The goal
%% of this project is to provide shared notebooks and a remote execution
%% environment for interpreted scripts.  \stephen{Are there significant
%%   differences involving versioning, provenance, annotation and other
%%   broad process support?}

%% {\bf bl.ocks, jsfiddle, plotly} and other web services for sharing code
%% and demos.

@@

Heer and Agarwala identify many design considerations for
collaborative visual analytics~\cite{Heer:2008:DCF} that
influenced our work.
RCloud notebooks, and the integrated version control system for them,
described in Section~\ref{sec:notebooks}, address modularity and granularity,
and artifact histories.
\emph{Starring}, the means for signaling interest in notebooks, described in
Section~\ref{sec:starring}, addresses social-psychological incentives,
recommendation, and voting and ranking. RCloud's integrated deployment
mechanism, described in Section~\ref{sec:deployment}, addresses the cost of
integration, content export, presentation and view sharing.

@@

There has been noteworthy work on specific techniques
to support collaborative or social code development and data analysis,
such as social bookmarking \cite{Millen:2006:DSB} \cite{Heer:2007:VAV}
and crowdsourcing \cite{Fast:2014:ECS}.
Similarly, there are computational methods to support high
performance execution in incremental code development
environments \cite{Guo:2010:TPI}.
The goal of RCloud is to define an environment in which many such
techniques may be integrated and made available to a broad community.
\begin{figure*}
\centering
\includegraphics[width=.95\linewidth]{fig/notebook/notebook.pdf}
\caption{\label{fig:notebook}An RCloud notebook is a sequence of
\emph{cells}, each a snippet of source in one of the supported languages (typically R, but Python and others are also supported) or Markdown. The main creation workflow involves editing notebooks, which are transparently stored as git repositories in GitHub, providing us with easy access primitives for version tracking. Notebooks can be executed as they're edited (left), or in a standalone viewer (right), via a slightly different URL. This lightweight provides a very low-friction mechanism for sharing results which we discuss in full in Section~\ref{sec:interviews}. }
\end{figure*}

%% \stephen{We have a problem we already went over some of this.}
%% Our work has been inspired by and benefited from other efforts
%% to improve data analysis environments and processes. Building in
%% the immense foundation of systems for statistical graphics,
%% recent work includes RStudio \cite{RStudio:2013:SWA},
%% the R packages Markdown \cite{Allaire:2014:MMR},
%% knitr \cite{Xie:2013:DDW}
%% and Shiny \cite{RStudio:2013:SWA},
%% and iPython notebooks \cite{Perez:2007:IAS}
%% to name a few. RStudio aims at providing an integrated development environment
%% for R, with support for publishing code in packages. Markdown,
%% knitR and Shiny augment R with sophisticated reporting capabilities, including
%% interactive web interfaces. IPython \cite{Perez:2007:IAS}
%% shares many of our goals, such as providing a comprehensive environment
%% for analysis and programming, and share-able documents on the web.
%% \stephen{Do we now need to introduce one of those remarks where we
%% explain that we are not the same as IPython.}

>>>>

Previous work has identified some requirements for supporting the
visual analytics process, including the work of multiple users.

\paragraph*{Social Data exploration and Analysis}
ManyEyes~\cite{Viegas:2007:MAS} was a landmark system for the
crowdsourced creation and publication of data
visualizations. Although ManyEyes supported a fixed set of
visual encodings, the system's success was a precursor to more
sophisticated solutions and an early indication that the world-wide
web was a suitable platform for data visualization. One of the main
challenges we faced in designing RCloud was providing an experience
for \emph{consuming} visualizations as seamless as ManyEyes's, while
sacrificing as little as possible on the generality of the analyses
themselves.

@@

\paragraph*{Provenance and versioning} As
mentioned, one central issue in exploratory analysis is that
problems can change quickly, often in the course of developing a
solution. As a result, systems should provide adequate support for
tracking \emph{changes} of the data analysis scripts. VisTrails was
one of the original systems for managing \emph{process provenance} ,
and demonstrated the value of capturing aspects of the processes that
surround data analysis experiments and tools, including detailed
history, collaboration, and deployment~\cite{Callahan:2006:VVM}.

@@

@@

Heer and Agarwala identify many design considerations for
collaborative visual analytics~\cite{Heer:2008:DCF} that
influenced our work.
RCloud notebooks, and the integrated version control system for them,
described in Section~\ref{sec:notebooks}, address {\it modularity and
granularity}, and {\it artifact histories}.
\emph{Starring}, the means for signaling interest in notebooks, described in
Section~\ref{sec:starring}, addresses social-psychological incentives,
recommendation, and voting and ranking. RCloud's integrated deployment
mechanism, described in Section~\ref{sec:deployment}, addresses the cost of
integration, content export, presentation and view sharing.

@@

There has been noteworthy work on specific techniques
to support collaborative or social code development and data analysis,
such as social bookmarking \cite{Millen:2006:DSB} \cite{Heer:2007:VAV}
and crowdsourcing \cite{Fast:2014:ECS}.
Similarly, there are computational methods to support high
performance execution in incremental code development
environments \cite{Guo:2010:TPI}.
The goal of RCloud is to define an environment in which many such
techniques may be integrated and made available to a broad community.
\begin{figure*}
\centering
\includegraphics[width=.95\linewidth]{fig/notebook/notebook.pdf}
\caption{\label{fig:notebook}An RCloud notebook is a sequence of
\emph{cells}, each a snippet of source in one of the supported languages (typically R, but Python and others are also supported) or Markdown. The main creation workflow involves editing notebooks, which are transparently stored as git repositories in GitHub, providing us with easy access primitives for version tracking. Notebooks can be executed as they're edited (left), or in a standalone viewer (right), via a slightly different URL. This provides a lightweight, low-friction mechanism for sharing results which we discuss in detail in Section~\ref{sec:interviews}. }
\end{figure*}
